---
title: "R Notebook"
output: html_notebook
---

## Packages

```{r message=FALSE, warning=FALSE}
library(devtools)
install_github("tranvivielodie/GpOutput2D/GpOutput2D")
install.packages("rlibkriging")

install_github("CharlieSire/FunQuant", upgrade="never")
```

```{r message=FALSE, warning=FALSE}
library(truncnorm)
library(ggplot2)
library(ggvoronoi)
library(FunQuant)
```

## Function to plot Voronoï diagram 

```{r}
plot_voronoi = function(prototypes, outputs= NULL,xlim = c(0,1), ylim =c(0,1)){
  outline = as.data.frame(rbind(c(xlim[1],ylim[1]),c(xlim[1],ylim[2]), c(xlim[2],ylim[2]), c(xlim[2],ylim[1])))
  protos = as.data.frame(do.call("rbind", lapply(prototypes, function(i){t(i)})))
  res = ggplot(data = protos) +
    stat_voronoi(geom="path", aes(x=V1, y=V2), outline = outline) +
    geom_point(aes(x=V1,y=V2), shape = 9, size = 5, col = "red") + theme_bw()
  if(!is.null(outputs)){
      outputs = as.data.frame(t(outputs))
      colnames(outputs) = c("V1", "V2")
      res = res + geom_point(data = outputs, aes(x = V1, y = V2)) 
  }
  return(res)
}

```

## Variables

We consider $X = (X_{1},X_{2}) \in \mathbb{R}^2$ a random input of a computer code $H$, with
$$\left\{
    \begin{array}{ll}
        X_{i} \sim \mathcal{N}_{t}(0,0.25^2, -1, 1), i=1,2 \\
        X_{1} \text{ and }X_{2}\text{ independent}
    \end{array}
\right.$$

where $\mathcal{N}_{t}(\mu,\sigma^2, a, b)$ is the Gaussian distribution of mean $\mu$, variance $\sigma^2$, truncated between $a$ and $b$.

```{r}
fX = function(x){
  return(
    dtruncnorm(x = x[1],mean = 0,sd = 0.25,a=-1, b=1)*dtruncnorm(x = x[2],mean = 0,sd = 0.25,a=-1, b=1))
}
```

The density function of $X$, denoted $f_{X}$, is represented below

```{r}
df_plot = as.data.frame(expand.grid(seq(-1,1,l=300), seq(-1,1,l=300)))
df_plot = cbind(df_plot, fX = apply(df_plot, 1, fX))
ggplot(df_plot,aes(x = Var1, y = Var2, z = fX)) + geom_contour_filled() + theme_bw()
                        
```

The computer code $H$ is defined with 

$$H(x) = \left\{
    \begin{array}{ll}
        (0,0) \text{ if } \lvert x_{1}\rvert \leq \alpha \\
        (\lvert x_{1} \rvert - \alpha, \lvert x_{2} \rvert) \text{ otherwise.}
    \end{array}
\right.$$

with $\alpha$ such that $P(H(X) = (0,0)) = 0.97.$

```{r}
ptrunc = -1+2*pnorm(1, 0,0.25)

alpha = qnorm(1.97*ptrunc/2, 0, 0.25)

H = function(x){
  if(abs(x[1]) < alpha){return(c(0,0))}
  else{return(c(abs(x[1])-alpha, abs(x[2])))}
}

```

## Classical Lloyd

This is the function to sample $X$ with density function $f_{X}$.

```{r}
sample_fX = function(n){cbind(rtruncnorm(n = n, mean = 0, sd = 0.25, a = -1, b = 1), rtruncnorm(n = n, mean = 0, sd = 0.25, a = -1, b = 1))}
```

Quantize $H(X)$ with $2000$ points, 3 starts and 5 cells

```{r}
set.seed(1)
inputs_fX = sample_fX(2000)
outputs_fX = apply(inputs_fX,1,H)
density_ratio_fX = rep(1, nrow(inputs_fX))
                                    
res_proto_fX = find_prototypes(data = outputs_fX,
                            nb_cells = 5,
                            multistart = 3,
                            density_ratio = density_ratio_fX)
```

```{r}
plot_voronoi(res_proto_fX$prototypes, outputs_fX, xlim = c(0, 1-alpha))

res_proto_fX$probas
```

## Std deviations 

With $1000$ bootstrap samples, we estimate the standard deviations of the estimators of the centroids of the optimal Voronoï cells

```{r}
set.seed(1)
std_centroid_fX = std_centroid(data = outputs_fX, 
              prototypes_list = list(res_proto_fX$prototypes),
              cells = 1:5,
              density_ratio = density_ratio_fX,
              bootstrap = 1000)

std_centroid_fX
```

And also the relative standard deviations of the estimators of the membership probabilities for the optimal Voronoï cells

```{r}
set.seed(1)
std_proba_fX = std_proba(data = outputs_fX, 
              prototypes_list = list(res_proto_fX$prototypes),
              cells = 1:5,
              density_ratio = density_ratio_fX,
              bootstrap = 3000)

std_proba_fX
```
## Uniform sampling

We introduce the uniform density $g$ and the function to sample $\tilde{X}$ with density function $g$.

```{r}
g = function(x){
  if(sum((x>-1)*(x<1))==2){return(1/4)}
  else{return(0)}
}

sample_g = function(n){ cbind (runif(n,-1,1), runif (n,-1,1))}
```

Quantize $H(X)$ with $2000$ points, 3 starts and 5 cells

```{r}
set.seed(1)
inputs_unif = sample_g(2000)
outputs_unif = apply(inputs_unif,1,H)
density_ratio_unif = compute_density_ratio(f = fX, 
                                      g = g, 
                                      inputs = inputs_unif)
                                    
res_proto_unif = find_prototypes(data = outputs_unif,
                            nb_cells = 5,
                            multistart = 3,
                            density_ratio = density_ratio_unif)
```


```{r}
plot_voronoi(res_proto_unif$prototypes, outputs_unif, xlim = c(0, 1- alpha))

res_proto_unif$probas
```

With $1000$ bootstrap samples, we estimate the standard deviations of the estimators of the centroids of the optimal Voronoï cells

```{r}
set.seed(1)

std_centroid_unif = std_centroid(data = outputs_unif, 
              prototypes_list = list(res_proto_unif$prototypes),
              cells = 1:5,
              density_ratio = density_ratio_unif,
              bootstrap = 1000)

std_centroid_unif
```

Comparison with the sample with $f_{X}$

```{r}

lapply(1:length(std_centroid_unif[[1]]), function(x){std_centroid_fX[[1]][[x]]/std_centroid_unif[[1]][[x]]})
```
And also the relative standard deviations of the estimators of the membership probabilities for the optimal Voronoï cells

```{r}
set.seed(1)

std_proba_unif = std_proba(data = outputs_unif, 
              prototypes_list = list(res_proto_unif$prototypes),
              cells = 1:5,
              density_ratio = density_ratio_unif,
              bootstrap = 3000)

std_proba_unif

```
Comparison with the sample with $f_{X}$. It highlights the fact that sampling with $f_{X}$ is much better for the first Voronoï cells, but not for the others.

```{r}
std_proba_fX[[1]]/std_proba_unif[[1]]

```


Example of the estimation of the standard deviation without bootstrapping

```{r}
large_sample_unif = sample_g(n=10^5)
large_outputs_unif = apply(large_sample_unif, 1, H)
large_density_ratio_unif = compute_density_ratio(f = fX, g = g, inputs = large_sample_unif)


std_proba_unif_large = std_proba(data = large_outputs_unif, 
              prototypes_list = list(res_proto_unif$prototypes),
              cells = 1:5,
              density_ratio = large_density_ratio_unif,
              nv = 2000)

std_proba_unif_large
```
## Hybrid sampling

As the uniform sampling reduces the variances for cells 2 to 5 but not for cell 1, an idea can be to sample with $f_{X}$ for cell 1 and with uniform sampling for the other cells.

Then we will create a list for these two samplings: the first element of the list is the sampling with $f_{X}$, and the second one the uniform sampling. 

The argument "sampling\_cells" indicates for each cell the associated index in the sampling list. 

```{r}
set.seed(1)
sampling_cells = c(1,2,2,2,2)

outputs_hybrid = list(outputs_fX, outputs_unif)
density_ratio_hybrid = list(density_ratio_fX, density_ratio_unif)

res_proto_hybrid = find_prototypes(data = outputs_hybrid,
                            nb_cells = 5,
                            multistart = 2,
                            density_ratio = density_ratio_hybrid,
                            method_IS = "percell",
                            sampling_cells = sampling_cells)



```
```{r}
plot_voronoi(res_proto_hybrid$prototypes, xlim = c(0, 1-alpha))

res_proto_hybrid$probas
```

```{r}
set.seed(1)

std_centroid_hybrid = lapply(1:5,function(i){
  std_centroid(prototypes_list = list(res_proto_hybrid$prototypes), 
  density_ratio = density_ratio_hybrid[[sampling_cells[i]]], 
  cells = i, 
  data = outputs_hybrid[[sampling_cells[i]]],
  bootstrap = 1000
  )})

std_centroid_hybrid
```



```{r}
set.seed(1)
std_proba_hybrid = unlist(sapply(1:5,function(i){
  std_proba(prototypes_list = list(res_proto_hybrid$prototypes), 
  density_ratio = density_ratio_hybrid[[sampling_cells[i]]], 
  cells = i, 
  data = outputs_hybrid[[sampling_cells[i]]],
  bootstrap = 3000
  )}))

std_proba_hybrid
```

## Evolving density 

We plot the limits of the Voronoi cells

```{r}
for(i in 1:5){
  plot(inputs_unif[res_proto_unif$cell_numbers ==i,1],inputs_unif[res_proto_unif$cell_numbers == i,2])
}
```

We define the following function

g\_adapt(x, cell, c) =
$$\left\{ 
    \begin{array}{ll}
        cste \times c \:\:\:\:\text{ if } abs(x_{1}) \in [min(abs(cell[,1])), max(abs(cell[,1]))] \: \& \: abs(x_{2}) \in [min(abs(cell[,2])), max(abs(cell[,2]))]\\
        cste \:\:\:\:\text{ otherwise.}
    \end{array}
\right.$$

And the function create\_sample to sample from g\_adapt

```{r}
g_adapt = function(x,cell,coeff){
  maxs = apply(abs(cell), 2, max)
  mins = apply(abs(cell),2,min)
  A = prod(maxs-mins)
  c1 = 1/(coeff*A+1-A)/4
  x = abs(x)
  if(sum((maxs-x)>0) == 2 & sum((x-mins)>0) == 2) {return(coeff*c1)}
  else{return(c1)}
}

create_sample = function(n, cell, coeff){
  maxs = apply(abs(cell), 2, max)
  mins = apply(abs(cell),2,min)
  A = prod(maxs-mins)

  prob1 = 1/(A*(coeff-1)+1)
  u = runif(n)
  
  tirage1 = matrix(runif(sum(u<prob1)*2)*2-1,ncol=2)
  tirage2 = cbind(sample(c(-1,1), size = sum(u>prob1), replace = TRUE)*(runif(sum(u>prob1))*(maxs[1]-mins[1])+mins[1]),sample(c(-1,1), size = sum(u>prob1), replace = TRUE)*(runif(sum(u>prob1))*(maxs[2]-mins[2])+mins[2]))
  return(as.data.frame(rbind(tirage1, tirage2)))
}

```

We then create a list density\_biased\_function with:

- The first element is fX
- The elements 2 to 5 is the function g\_adapt(x, cell, 10^3)

And we create a list of functions to generate 2000 inputs from the densities in density\_biased\_function

```{r}
density_biased_function = c(list(function(x, cell){fX(x)}), lapply(1:4, function(i){function(x, cell){g_adapt(x, cell, 10^3)}}))

samp_cell1 = function(cell){
  set.seed(10)
  sample_fX(2000)
}

inputs_function = c(list(samp_cell1), lapply(1:4, 
  function(i){function(cell){
    set.seed(10)
    create_sample(2000,cell, 10^3)}}))

res_proto_pred_adapt = find_prototypes(
  starting_proto = res_proto_hybrid$prototypes, 
  method_IS = "percell", 
  density_ratio = density_adapt, 
  sampling_cells = 1:5,
  inputs_ref = inputs_unif,
  data_ref = outputs_unif,
  density_function = fX,
  density_biased_function = density_biased_function,
  inputs_function = inputs_function,
  outputs_function = function(df){apply(df,1,H)},
  print_progress = TRUE,
  threshold = 0.01)

```
Finally, we will assess the standard deviations for the optimal prototypes

We then need to reproduce the last samplings of the algorithm. 

We get matrices of the points inside the Voronoi cells (here stored in cells\_reff), and then to generate inputs from these matrices, following the density density\_ratio\_adapt(x, cell).



```{r}
numbers_ref = get_cell_numbers(outputs_unif,res_proto_pred_adapt$prototypes)

cells_ref = lapply(1:5, function(h){as.data.frame(inputs_unif[numbers_ref == h,])}) ## get the

inputs_list = lapply(1:5, function(h){inputs_function[[h]](cells_ref[[h]])})

outputs_adapt = lapply(1:5, function(h){apply(inputs_list[[h]],1,H)})

density_ratio_adapt = lapply(1:5, function(h){compute_density_ratio(fX, function(x){density_biased_function[[h]](x, cells_ref[[h]])}, inputs_list[[h]])})
```

```{r}
set.seed(1)
std_centroid_adapt = lapply(1:5,function(i){
  std_centroid(prototypes_list = list(res_proto_pred_adapt$prototypes), 
  density_ratio = density_ratio_adapt[[i]], 
  cells = i, 
  data = outputs_adapt[[i]],
  bootstrap = 1000
  )})

std_centroid_adapt

```

```{r}
lapply(1:length(std_centroid_unif[[1]]), function(x){std_centroid_adapt[[x]][[1]][[1]]/std_centroid_hybrid[[x]][[1]][[1]]})
```


```{r}
set.seed(1)
std_proba_adapt = unlist(sapply(1:5,function(i){
  std_proba(prototypes_list = list(res_proto_pred_adapt$prototypes), 
  density_ratio = density_ratio_adapt[[i]], 
  cells = i, 
  data = outputs_adapt[[i]],
  bootstrap = 3000
  )}))

std_proba_adapt

```
Comparison with hybrid sampling and sampling with $f_{X}$
```{r}
std_proba_adapt/std_proba_hybrid
std_proba_adapt/std_proba_fX[[1]]

```
The reduction of the variances for the cells 2 to 5 is significative. 
